---
title: "Short analysis on SV types detected in NA12878 and Exceptional Responders"
output: html_notebook
---
# NA12878

Use the script "SV_type.R" in the main folder (~/../NA12878_benchmark) to get the SV types detected by each caller from NA12878 into one data frame ("svt").

That script is also replicated below:
```{r Generate sv type data, include=F, eval=F}
# This script imports SV type data from the BND converter (See SvABA_BND_annotation.R)
# to svaba_svgr objects
# and then analyse the type of SVs detected in each caller 
# using NA12878 at different coverage levels

setwd("~/GitRepo/summerproject2021/NA12878_benchmark/SvABA_processing")

# BND from csv to svaba_svgr by 
med.svaba.uniq=read.csv("med_svaba_converted.csv")
ft.svaba.uniq=read.csv("50pc_converted.csv")
large.svaba.uniq=read.csv("svaba_converted.csv")

# svgr#elementMetadata$sourceID ~ csv$ID
# svgr@elementMetadata$svtype ~ csv$SV_type

med_svaba_svgr@elementMetadata$svtype[which(med_svaba_svgr@elementMetadata$sourceId == med.svaba.uniq$ID)]=med.svaba.uniq$SV_type

ft_svaba_svgr@elementMetadata$svtype[which(ft_svaba_svgr@elementMetadata$sourceId == ft.svaba.uniq$ID)]=ft.svaba.uniq$SV_type

svaba_svgr@elementMetadata$svtype[which(svaba_svgr@elementMetadata$sourceId == large.svaba.uniq$ID)]=large.svaba.uniq$SV_type

# sv type
dflarge=large_svgr@elementMetadata[, c("svtype", "Caller")]
dfft=ft_svgr@elementMetadata[, c("svtype", "Caller")]
dfmed=med_svgr@elementMetadata[, c("svtype", "Caller")]

# cov
dflarge$Cov="35X"
dfft$Cov="18X"
dfmed$Cov="6X"


dflarge=as.data.frame(dflarge)
dfft=as.data.frame(dfft)
dfmed=as.data.frame(dfmed)

# join
plyr::join(dflarge, dfft, type="full") %>% plyr::join(dfmed, type="full") -> svt

# plot
ggplot(svt) + aes(x=svtype, fill=Caller) + geom_bar(position="dodge") + facet_wrap(~Cov) + theme_bw()
ggsave("~/GitRepo/summerproject2021/NA12878_benchmark/Figures/SV_type_by_coverage.png")

# reorder panel header level for PR plot
svt$Cov <- factor(svt$Cov, levels = c("35X", "18X", "6X"))
```


```{r Plot SV type by caller}
# plot
ggplot(svt) + aes(x=svtype, fill=Caller) + geom_bar(position="dodge") + facet_wrap(~Cov, nrow=1) + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1.2 ,vjust=1.2)) + ylab("Count") + xlab("Type of structural variant")
ggsave("SV_type_by_caller.png")
```

```{r Plot SV type by coverage}
# sv type by coverage
ggplot(svt) + aes(x=svtype, fill=Cov) + geom_bar(position="dodge") + theme_bw() +
  ylab("Count") + xlab("Type of structural variant") + 
  paletteer::scale_fill_paletteer_d("LaCroixColoR::PommeBaya", name="Coverage")
ggsave("SV_type_by_coverage.png")
```

# Exceptional Responders

You'll need the data frame "ExRes_q", which is the qualified call set from the six exceptional responders' germline samples.

The data were generated in "~/../Exceptional_responders/ExRes_benchmark.Rmd". We will import it here for your convenience:
```{r Import ExRes variant data}
ExRes_q=read_rds("~/GitRepo/summerproject2021/Exceptional_responders/ExRes_q.rds")
```

```{r Plot SV type in the Exceptional Responders Cohort}
# SV type and length, but must wrap by caller (or else duplicate counts confound)
ExRes_q %>% 
  filter(!is.na(SV_length)) %>%
  ggplot() + aes(x=abs(SV_length), color = SV_type) + 
  geom_line(aes(y=..count..), stat="bin", size = 1.05) +
  facet_wrap(~Caller) + 
  theme_classic() + 
  paletteer::scale_colour_paletteer_d("khroma::bright", name = "SV type") +
  scale_x_log10(name = "SV size (log-scaled)", 
                breaks = c(1, 20, 1e2, 1e3, 1e4, 1e5, 1e6, 1e7, 1e8), 
                label = c("1 bp", "20 bp", "100 bp", "1 kb", "10 kb", "100 kb", "1 Mb", "10 Mb", "100 Mb")) +
  scale_y_continuous(trans = "log10", 
                name = "Counts (log-scaled)",
                breaks= c(10, 10^2, 10^3, 10^4, 10^5), 
                label = c("10", "100", "1k", "10k", "100k"))

ggsave("SV_length_trend.png")
```


