---
title: "SvABA output processing"
output: html_notebook
---
Below is a script to remove the three redundant columns in the raw VCF files from SvABA output.

**IMPORTANT** Be sure to change the working directory to where the VCF file is located to prevent errors.


# 1. Use Bash to rename the columns of interest
```{bash Get column header}
# this is the raw vcf file from svaba

# copy the line after "FORMAT"
# /cromwell_root/.../sample_name.bam

grep "#CHROM" NA12878.svaba.sv.vcf
```


```{bash Rename columns of interest}
# ... tFORMAT\t/***PASTE HERE***...../(including)sample_name.bam"
# change this: tA\tB\tC\t***sample_name*** > sample_name.sv.vcf
# rename the output if you want to retain the original vcf file

cat NA12878.svaba.sv.vcf | awk '{gsub("FILTER\tINFO\tFORMAT\t/cromwell_root/gatk-test-data/wgs_bam/NA12878_24RG_hg38/NA12878_24RG.hg38.bam", "FILTER\tINFO\tFORMAT\tA\tB\tC\tNA12878")}1' > NA12878.svaba.sv.vcf
```

The columns are renamed accordingly and the original VCF files overwritten.


# 2. Use R to remove redundant columns
```{r Process vcf file in R}
# the three columns have been renamed but not yet removed
library(vcfR)

to.be.cleaned.vcf <- read.vcfR("NA12878.svaba.sv.vcf")

# remove the 3 redundant columns
  #dim(to.be.cleaned.vcf@gt[, c(2:4)])
to.be.cleaned.vcf@gt = to.be.cleaned.vcf@gt[, -c(2:4)]
  #dim(to.be.cleaned.vcf@gt)

# write the cleaned object to a new vcf file
write.vcf(to.be.cleaned.vcf, "NA12878.svaba.vcf") ## note that the .sv was removed in the new file
```

**IMPORTANT**
Even if the original input VCF was decompressed, the resultant file needs to be decompressed again!

Rename output (the one without .sv) as *.vcf.gz and open to decompress (Worked on MacOS).

```{r After processed vcf is decompressed test read in}
# test output file
# should be able to load directly 
# can also check total SV calls
test.vcf <- read.vcfR("NA12878.svaba.vcf")
```

